SMT-Based Generation of Symbolic Automata \thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2019-03-30 2017-Xudong-SMT/ActaInformatica.tex

%\TODO{Formalize the proof}
\begin{proof}
In Theorem 2, we have already proved the result automata $OA_1$ and $OA_2$ of the two algorithm are the same without the assignments.

Define the function $States$ as before, and  $Assignments : \langle OT\rangle\times\langle State\rangle\rightarrow\langle Assign\rangle$ the function
computing the set of assignments of a global state from a given set of OTs.

The structure implicitely computed by the algorithm is a triple $ Aut= \langle OTs, States, Assigns\rangle$.

Then the assignments set of $OA_1$ is $Assignments(L1, States(L1))$ and that of $OA_2$ is $Assignments(L2_{sat}, States(L2_{sat}))$. The computed triples are 
\[Aut_1 = \langle L1_{sat}, States(L1_{sat}), Assignments(L1, States(L1))\rangle\] and \[Aut_2 = \langle L2_{sat}, States(L2_{sat}), Assignments(L2_{sat}, States(L2_{sat}))\rangle \ . \]

Based on the notation of the triple, we define the function $Refine : Aut \rightarrow Aut$ refining the result automaton together with the assignments,
\begin{align*}
Refine(Aut) &=\langle OT', States(OT'), Assignments(OT', States(OT'))\rangle \\
&=Aut'
\end{align*}
where $OT' = OT - OT_{unsat}$, $OT_{unsat}$ is the set of transitions checked unsatisfiable in this iteration.

Given that every iteration will remove some unsatisfiable transitions, $Refine$ is a monotone function. There exists a sequence $Aut^n = Refine(Aut^{n-1})$ s.t. $|Aut^n| < |Aut^{n-1}|$.
Let $Aut''$ contains the $OT''$ which has no more unsatisfiable transition, then there is $Aut'' = Refine(Aut'')$ and $|Aut''| = |Refine(Aut'')|$.
Let $Aut''_1$ and $Aut''_2$ be the two
 fixed points with $Refine(Aut''_1) = Aut''_1$, $Refine(Aut''_2) = Aut''_2$.

For the transition sets in $Aut''_1$ and $Aut''_2$, $L1''_{sat}=L1_{reach}$ and $L2''_{sat}=L2_{reach}$, we need to check their equivalence.
At the first iteration, the unsatisfiable transitions in $OA_1$ has only one case according to the relation between the assertions to check,
\begin{align*}
&\quad Pred\wedge \bigvee_i Assignments_i(L1, States(L1) \text{ is false}\\
&\quad\longrightarrow Pred\wedge \bigvee_j Assignments_j(L2_{sat}, States(L2_{sat})\text{ is false}\\
&\Rightarrow \forall t\in L1_{unsat} \longrightarrow t\in L2_{unsat}\\
&\Rightarrow L1_{unsat}\subseteq L2_{unsat}.
\end{align*}
However, its reverse direction is not fulfilled. Recall that $L1_{sat}=L2_{sat}$, these relations can imply that,
\begin{align*}
Assignments(L1, &States(L1))\supseteq Assignments(L2_{sat}, States(L2_{sat}))\\
&\supseteq Assignments(L1_{sat}-L1_{unsat}, States(L1_{sat}-L1_{unsat}))\\
&\supseteq Assignments(L2_{sat}-L2_{unsat}, States(L2_{sat}-L2_{unsat}))
\end{align*}
So after the first iteration, the unsatisfiable transitions in $OA_1$ has two cases,
\begin{align*}
&\forall t\in L1'_{unsat} \longrightarrow t\in L2'_{unsat}\\
\text{ or }&\forall t\in L1'_{unsat} \longrightarrow t\notin L2'_{sat}.
\end{align*}
By induction, after removing all the transitions not in $L2''_{sat}$, we obtain  $L1_{reach}=L2_{reach}$, which implies that $States_{L1_{reach}}=States_{L2_{reach}}$ and
$$Assignments\\(L1_{reach},States_{L1_{reach}})=Assignments(L2_{reach},States_{L2_{reach}})\ .$$
\hfill\qed
\end{proof}

>>>>

%\TODO{Formalize the proof}
\begin{proof}
In Theorem 2, we have already proved that the result automata $OA_1$ and $OA_2$ of two algorithms are the same if the pNet contains no assignment to its state variables.

We define the function $States$ as before, and the function $Assignments : \langle OT\rangle\times State\rightarrow\langle Assign\rangle$
computing the set of assignments of a global state from a given set of OTs.

The structure implicitly computed by the algorithm is a triple $ Aut= \langle OTs, States,\\ Assigns\rangle$.

Then the assignments set of $OA_1$ is $\{Assignments(L1, State_{i})\}^{State_{i}\in States(L1)}$ and $OA_2$ is $\{Assignments(L2_{sat}, State_{j})\}^{State_{j}\in States(L2_{sat})}$. The result triples should be 
\[Aut_1 = \langle L1_{sat}, States(L1_{sat}), \{Assignments(L1, State_{i})\}^{State_{i}\in States(L1)}\rangle\] and \[Aut_2 = \langle L2_{sat}, States(L2_{sat}), \{Assignments(L2_{sat}, State_{j})\}^{State_{j}\in States(L2_{sat})}\rangle \ . \]

Basing on the notation of the triple, we define the function $Refine : Aut \rightarrow Aut$ implementing $makeReachableSubAutomaton(S0,OTs)$, it uses the assignments of the previous step to refine the computed reachable space.
\begin{align*}
Refine(Aut) &=\langle L', States(L'), \{Assignments(L', State_{i})\}^{State_{i}\in States(L')}\rangle \\
&=Aut'
\end{align*}
where $L' = L - L_{unsat}$, $L_{unsat}$ is the set of unsatisfiable transitions checked in this iteration. So there must have $|L'|<|L|$, and potentially, their target states and corresponding assignments could be reduced: 
\[States_{L'}=\{State_{init}\}\cup \bigcup States(L'),\]
\[Assignments_{L',States_{L'}}=\bigcup\{Assignments(L', State_{i})\}^{State_{i}\in States_{L'}} \]
\[s.t. |States_{L'}|\le|States_{L}|\wedge|Assignments_{L',States_{L'}}|\le|Assignments_{L,States_{L}}|.\]

In one single iteration, satisfiability of a transition $\ostate{\tilde{t}}\xrightarrow{l}\ostate{\tilde{t'}}\in L$ is checked through whether 
\[\bigvee_{Assign_i\in Assignments_{L, States_{L}}}(Pred_{l}\wedge Assign_{i})\text{ is true}.\]
More transitions could be unsatisfiable after the reduction of the assignments set, so the function need iterate several times. Given that in each iteration, the number of triples decreases, so $Refine$ is a monotone function. And there exists a sequence \[Aut^n = Refine(Aut^{n-1})\quad s.t. |Aut^n| < |Aut^{n-1}|.\]
The function terminates if $Aut''$ contains no more unsatisfiable transition, then there is a fixed-point $Aut'' = Refine(Aut'')$ and $|Aut''| = |Refine(Aut'')|$.

Let $Aut''_1$ and $Aut''_2$ be two fixed-points that $Refine(Aut''_1) = Aut''_1$, $Refine(Aut''_2) = Aut''_2$. Focusing on the transition sets in $Aut''_1$ and $Aut''_2$, $L1''_{sat}=L1_{reach}$ and $L2''_{sat}=L2_{reach}$, we need to check their equivalence.

First, there is an inclusion relation between the initial assignments sets, 
\begin{align*}
&L2_{sat}\subseteq L1\Rightarrow States_{L2_{sat}}\subseteq State_{L1}\\
\Rightarrow &Assignments_{L2_{sat}, States_{L2_{sat}}}\subseteq Assignments_{L1, States_{L1}}
\end{align*}

Then we come to see what changes have occurred during the iterations. In the first iteration, the inclusion relation between their assignments implies an inclusion relation of unsatisfiable transitions. If there are two transitions $\ostate{\tilde{t}}\xrightarrow{l}\ostate{\tilde{t'}} \in L1_{unsat}$ and $\ostate{\tilde{t}}\xrightarrow{l}\ostate{\tilde{t'}} \in L2_{unsat},$ then
\begin{align*}
&\bigvee_{Assign_i\in Assignments_{L1, States_{L1}}} (Pred_{l}\wedge Assign_i)\text{ is false}\\
\longrightarrow&\bigvee_{Assign_j\in Assignments_{L2_{sat}, States_{L2_{sat}}}} (Pred_{l}\wedge Assign_j)\text{ is false}.
\end{align*}
The relation that $\forall\ostate{\tilde{t}}\xrightarrow{l}\ostate{\tilde{t'}} \in L1_{unsat},\exists\ostate{\tilde{t}}\xrightarrow{l}\ostate{\tilde{t'}} \in L2_{unsat}$, implies that $L1_{unsat}\subseteq L2_{unsat}.$

However, its reverse direction is not fulfilled. Recall that $L1_{sat}=L2_{sat}$, more inclusion relations of the assignments can be implied,
\begin{align*}
Assignments_{L1, States_{L1}}&\supseteq Assignments_{L2_{sat}, States_{L2_{sat}}}\\
&\supseteq Assignments_{L1_{sat}-L1_{unsat}, States_{(L1_{sat}-L1_{unsat})}}\\
&\supseteq Assignments_{L2_{sat}-L2_{unsat}, States_{(L2_{sat}-L2_{unsat})}}
\end{align*}

Let $L1'_{sat}=L1_{sat}-L1_{unsat}$ and $L2'_{sat}=L2_{sat}-L2_{unsat}$
Then in the next iteration, the unsatisfiable transition $\ostate{\tilde{t}}\xrightarrow{l}\ostate{\tilde{t'}}$ removed from $L1'_{sat}$ might be the same transition (up to renaming) removed from $L2'_{sat}$. In this case,
\[\ostate{\tilde{t}}\xrightarrow{l}\ostate{\tilde{t'}}\in L1'_{unsat} \longrightarrow \ostate{\tilde{t}}\xrightarrow{l}\ostate{\tilde{t'}}\in L2'_{unsat}.\]
Or it is the transition has not removed in the last iteration because of 
\[Assignments_{L2_{sat}, States_{L2_{sat}}}\subseteq Assignments_{L1, States_{L1}}.\] 
We have already proved that 
\[Assignments_{L1'_{sat}, States_{L1'_{sat}}}\subseteq Assignments_{L2_{sat}, States_{L2_{sat}}},\]
so those transitions will be removed this time. In this case,
\[\ostate{\tilde{t}}\xrightarrow{l}\ostate{\tilde{t'}}\in L1'_{unsat} \longrightarrow \ostate{\tilde{t}}\xrightarrow{l}\ostate{\tilde{t'}}\notin L2_{sat}.\]

By induction, if there is no more unsatisfiable transitions in $L2''_{sat}$, $L1''_{sat}$ continues removing all the transitions not in $L2''_{sat}$, until $L1_{reach}=L2_{reach}$. It implies that $States_{L1_{reach}}=States_{L2_{reach}}$ and
$$Assignments_{L1_{reach},States_{L1_{reach}}}=Assignments_{L2_{reach},States_{L2_{reach}}}.$$
\hfill\qed
\end{proof}
